---
title: "DS Final Project"
author: "Zachary Dougherty"
date: "November 29, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading the data}
season_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\nba-players-stats\\Seasons_Stats.csv"
,
  header = T,
  sep = ",")
#P1
players<- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\nba-players-stats\\Players.csv", sep = "," )
Player_data<- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\nba-players-stats\\player_data.csv",
  head = TRUE,
  sep = ",")
GSW_stats18<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW_statistics.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats18) <- c("team(18)", "team/G(18)", "L Rank(18)")
GSW_stats17<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2016-2017.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats17) <- c("team(17)", "team/G(17)", "L Rank(17)")
GSW_stats16<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2015-2016.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats16) <- c("team(16)", "team/G(16)", "L Rank(16)")
GSW_stats15<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2014-2015.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats15) <- c("team(15)", "team/G(15)", "L Rank(15)")
GSW_stats14<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2013-2014.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats14) <- c("team(14)", "team/G(14)", "L Rank(14)")
GSW_stats13<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2012-2013.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats13) <- c("team(13)", "team/G(13)", "L Rank(13)")
GSW_stats12<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2011-2012.csv",
                       header = TRUE,
                       sep = ",")
row.names(GSW_stats12) <- c("team(12)", "team/G(12)", "L Rank(12)")
GSW_stats11<- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\GSW2010-2011.csv",
                       header = TRUE,
                       sep = ",")
#P2
twenty_ten_eleven_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2010_11_team_per_game_stats.csv",
  sep = ",")
twenty_eleven_twelve_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2011_12_team_per_game_stats.csv",
  sep = ",")
twenty_twelve_thirteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2012_13_team_stats_per_game.csv",
  sep = ",")
twenty_thirteen_fourteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2013_14_team_per_game_stats.csv",
  sep = ",")
twenty_fourteen_fifteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2014_15_team_stats_per_game.csv",
  sep = ",")
twenty_fifteen_sixteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2015_16_team_stats_per_game.csv",
  sep = ",")
twenty_sixteen_seventeen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2016_17_team_per_game_stats.csv",
  sep = ",")
twenty_seventeen_eighteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2017_18_team_stats_per_game.csv",
  sep = ",")
twenty_eighteen_nineteen_team_stats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\2018_19_season_stats_per_game.csv",
  sep = ",")
#P3
d_rose <- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\D_Rose_stats.csv",
                   sep = ",")
lbj <- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\LBJ_stats.csv",
                   sep = ",")
kd <- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\KD_stats.csv",
                   sep = ",")
sc <- read.csv("C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\Steph_Curry_stats.csv",
                   sep = ",")
russ_wb <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\Russell_Westbrook_stats.csv",
                   sep = ",")
james_h <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\James_Harden_stats.csv",
                   sep = ",")
#P4
nbaStats <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\nbastats.csv") 
nbaRatings <- read.csv(
  "C:\\Users\\bbygbbyg\\Documents\\DS_Final_Project\\nbaratings.csv") 

library(RColorBrewer)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
```

#Introduction
From the data collected from the GSW stats we were interested to see what factors separate the
GSW from other teams, what makes them so much more efficient and better than others. To do so we
broke it up the statistics so that we can notice a specific thing that GS is doing right that
others can't do. 

#Part 1
The average person knows that GS is known for shooting and making multiple 3 pointers. The
starting five players on the warriors are all efficient when it comes to making field goals
behind the arc. To our surprise, the data displays says that GS doesn't even take the most shots
in the NBA. They are ranked 17 in the league for 3 pointers attempted in a season. They are
ranked 17 but yet they have attempted 2,370 in the 2017-2018 season. GS isn't even ranked top 10
for 3 pointers attempted but yet they are number 1 for 3 point percentage. Their efficiency on
the court for making 3 pointers has forced other teams to shoot more shots behind the arc to keep
up, which often times results in their defeat. By putting more stress on other teams to catch up
teams become  desperate to make up the point difference. The team also lead the league with the
highest average of points per game, 113.5. If you look back decades or even back in the 2000s,
teams hardly scored over 100 points per game. When it did happen it was because the team was on
fire making on their shot. The team has transformed the league by forcing others to keep up with
their high scoring, resulting in teams taking more 3s than GS themselves. GS overall is a high
scoring machine, they also lead the league in 2 point percentages while once again not being the
top 10 team in 2 pointers attempted. They were ranked 20 in the league, even lower in rank
compared to 3 pointers, and yet they make more than anyone else. The effectiveness of the team is
seen through their ball movement and how they have set up their offense to always have options.
```{r part 1}
row.names(GSW_stats11) <- c("team(11)", "team/G(11)", "L Rank(11)")
GSW_stats<- rbind(GSW_stats11,GSW_stats12,GSW_stats13,GSW_stats14,
                  GSW_stats15,GSW_stats16,GSW_stats17,GSW_stats18)
season<- data.frame(seasons=1:8)
threePA11<- GSW_stats11$X3PA[1]
threePA12<- GSW_stats12$X3PA[1]
threePA13<- GSW_stats13$X3PA[1]
threePA14<- GSW_stats14$X3PA[1]
threePA15<- GSW_stats15$X3PA[1]
threePA16<- GSW_stats16$X3PA[1]
threePA17<- GSW_stats17$X3PA[1]
threePA18<- GSW_stats18$X3PA[1]
threePA<- rbind(threePA11,threePA12,threePA13,threePA14,threePA15,threePA16,threePA17,threePA18)
threePA_season<- cbind(season,threePA)

threePA_graph<- ggplot(threePA_season,
                       aes(x=seasons, y=threePA_season$threePA,fill=seasons)) +
  geom_bar(stat = "identity",fill="#e2988b",color="black") +
  xlab("Season")+
  ylab("3 Ppointers Attempted")+
  ggtitle("GSW 3's Attempted Per Season") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))
threePA_graph

threeP11<- GSW_stats11$X3P[1]
threeP12<- GSW_stats12$X3P[1]
threeP13<- GSW_stats13$X3P[1]
threeP14<- GSW_stats14$X3P[1]
threeP15<- GSW_stats15$X3P[1]
threeP16<- GSW_stats16$X3P[1]
threeP17<- GSW_stats17$X3P[1]
threeP18<- GSW_stats18$X3P[1]
threeP<- rbind(threeP11,threeP12,threeP13,threeP14,threeP15,threeP16,threeP17,threeP18)
threeP_season<- cbind(season, threeP)

threeP_graph<- ggplot(threeP_season, aes(x= threeP_season$seasons, y=threeP_season$threeP))+
  geom_bar(stat = "identity", color = "black",fill="#004966") +
  xlab("Season") +
  ylab("3 Pointers Made") +
  ggtitle("GSW 3's Made Per Season") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))
threeP_graph
```
From what is displayed by the graphs you can see nearly see a correlation between the amount of
3s attempted and the percentage of 3s made to the success of the team. For the past 8 seasons you
can see that the Warriors have adopted the three point shot to be their main attribute as a team.
Out of the 8 seasons the warriors have lead the league in 3 point percentage four times, being
top 5 in the league all 8 years. They have only lead 3 point attempts once in the past 8 seasons
and that season they broke the record for the amount of wins in a season. You could say practice
makes perfect because from 2010-2013 the team never broke 2000 3 pointers in a season. It was
after the 2012-2013 that the team would have stay about 2000 for the next 5-6 years. Out of those
5-6 years the team won 3 championships and have made it to playoffs every single year since. 

#Part 2
After analyzing Golden State's trends during the past nine seasons, we wanted to see how the rest
of the league matched Golden State's trends. We decided to look at the 3 point attempts and makes
and 2 point attempts and makes of all other teams besides Golden State for the past nine
seasons.
```{r part 2}
#All stats for seasons 2010-11 to 2018-19
all <- rbind(twenty_ten_eleven_team_stats,
             twenty_eleven_twelve_team_stats,
             twenty_twelve_thirteen_team_stats,
             twenty_thirteen_fourteen_team_stats,
             twenty_fourteen_fifteen_team_stats,
             twenty_fifteen_sixteen_team_stats,
             twenty_sixteen_seventeen_team_stats,
             twenty_seventeen_eighteen_team_stats,
             twenty_eighteen_nineteen_team_stats)
#Want to check rest of the league without Golden State
all <- filter(all,Team!="Golden State Warriors"&Team!="Golden State Warriors*")
#Some plots
three_att_box <- ggplot(all,aes(x=as.factor(Season),y=X3PA,color=Season)) +
  geom_boxplot(fill="black") +
  xlab("Season") +
  ylab("Three Point Attempts") +
  ggtitle("Three Point Attempts all Teams") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

three_made_box <- ggplot(all,aes(x=as.factor(Season),y=X3P,color=Season)) +
  geom_boxplot(fill="black") +
  xlab("Season") +
  ylab("Three Point Made") +
  ggtitle("Three Point Made all Teams") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

two_att_box <- ggplot(all,aes(x=as.factor(Season),y=X2PA,color=Season)) +
  geom_boxplot(fill="black") +
  xlab("Season") +
  ylab("Two Point Attempts") +
  ggtitle("Two Point Attempts all Teams") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

two_made_box <- ggplot(all,aes(x=as.factor(Season),y=X2P,color=Season)) +
  geom_boxplot(fill="black") +
  xlab("Season") +
  ylab("Two Point Made") +
  ggtitle("Two Point Made all Teams") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

three_att_box
three_made_box
two_att_box
two_made_box
```
From these boxplots, we can see that the average 3 point attempts and makes has increased
steadily even before Golden State's exploitation of the shot. However, after the 2014-15 season,
we can see multiple outliers above the mean. For instance, during the 2016-17 season, one team
was attempting nearly 40 3 point shots per game, about 15 more than the average. The next two
season have outliers nearing the 45 attempts mark. Clearly, certain teams have tried to emphasize
the 3 point shot in their gameplan, and the high number of 3 point shots made speaks to the
efficacy of this shift. The number of 2 point attempts has slightly decreased over the past nine
seasons, with many low outliers, some being almost 20 attempts below averages. These plots may
show that a general trend in the league is occurring, with a few teams drastically altering their
game plan in an attempt to emulate the success of the Golden State Warriors.

#Part 3
Though team trends are important to observe, teams are influenced by the players on them,
and so it is necessary for us to analyze how individual players have changed their game as well.
We chose all the players that had either won MVP, were the scoring champion or led the league in
win shares in the past nine seasons. Win shares is a complex statistic that essentially rates how
many wins on a team can be attributed to a particular player on that team, it is a measure of how
important that player was to earning wins for the team. There were only six players that shared
all these different awards over the past seasons, LeBron James, James Harden, Derrick Rose, Kevin
Durant, Stephen Curry, and Russell Westbrook.
```{r part 3.1}
#Using league MVP's, Pt champions, and WS champions

#Adding name columns
d_rose <- cbind(d_rose,"Name" = rep("Derrick Rose"))
lbj <- cbind(lbj,"Name" = rep("LeBron James"))
kd <- cbind(kd,"Name" = rep("Kevin Durant"))
sc <- cbind(sc,"Name" = rep("Stephen Curry"))
russ_wb <- cbind(russ_wb,"Name" = rep("Russell Westbrook"))
james_h <- cbind(james_h,"Name" = rep("James Harden"))
#Combining all top players statistics
top_players <- rbind(d_rose,lbj,kd,sc,russ_wb,james_h)

#Plots
top_players_three_att <- ggplot(top_players,
                              aes(x=interaction(Season,Name),
                                  y=X3PA,fill=Name)) +
  geom_bar(stat = "identity",color = "black") +
  xlab("Interaction of Season and Player") +
  ylab("Three Point Attempt") +
  scale_fill_brewer(palette = "YlGn") +
  ggtitle("Three Point Attempts of Top Players") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

top_players_three_made <- ggplot(top_players,
                              aes(x=interaction(Season,Name),
                                  y=X3P,fill=Name)) +
  geom_bar(stat = "identity",color = "black") +
  xlab("Interaction of Season and Player") +
  ylab("Three Point Made") +
  scale_fill_brewer(palette = "OrRd") +
  ggtitle("Three Point Shots Made by Top Players") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

top_players_three_att
top_players_three_made
```
It is important to keep in mind exactly what we are trying to look for in these plots. We want to
see if players who are performing at the highest possible level are utilizing the three point 
shot more during games. We can see that three players truly stand out from the rest, those being
Stephen Curry, Russell Westbrook, and James Harden. These three players have all attempted and
made more 3 point field goals than their peers on the plot, and all three of these players have
been either MVP, scoring champion, or leader in win shares during the past three seasons. On the
opposite extreme, we see a player like Derrick Rose who, after getting injured and losing some of
his athleticism and playing ability, fell from utilizing the 3 point shot. But, during his past
two seasons with Minnesota, he has gotten to a point where this season, he is making the most 3
point shots per game in his entire career, and is considered to be having an NBA comeback. LeBron
James has also increased his 3 point attempts and makes per game, and is considered the best
player in the NBA today.

We also wanted to examine if the roles of certain positions changed with league wide trends. We
wanted to focus on the center, as the center is typically a position played inside the key and
doesn't traditionally come out past the three point line looking for a shot. We wanted to see
whether big men are also joining this trend and compare that to the 3 point attempts and makes of
point guards, a position which traditionally does take three point shots. We decided to only
include players who played at least 50 games a season and 25 minutes per game, otherwise players
that have little minutes played could skew the data.
```{r part 3.2}
season_stats_post_2010 <- subset(season_stats,Year>2009)

#Positions
#Have to play at least 50 G
pgs <- filter(season_stats_post_2010,Pos=="PG"&Tm!="TOT"&G>50&MP>25)

PG2010 <- summarise_all(filter(pgs,Year==2010),funs(mean))
PG2011 <- summarise_all(filter(pgs,Year==2011),funs(mean))
PG2012 <- summarise_all(filter(pgs,Year==2012),funs(mean))
PG2013 <- summarise_all(filter(pgs,Year==2013),funs(mean))
PG2014 <- summarise_all(filter(pgs,Year==2014),funs(mean))
PG2015 <- summarise_all(filter(pgs,Year==2015),funs(mean))
PG2016 <- summarise_all(filter(pgs,Year==2016),funs(mean))
PG2017 <- summarise_all(filter(pgs,Year==2017),funs(mean))
pgs_avgs <- rbind(PG2010,PG2011,PG2012,PG2013,PG2014,PG2015,PG2016,PG2017)
pgs_avgs$Pos <- rep("PG")

centers <- filter(season_stats_post_2010,Pos=="C"&Tm!="TOT"&G>50&MP>25)

C2010 <- summarise_all(filter(centers,Year==2010),funs(mean))
C2011 <- summarise_all(filter(centers,Year==2011),funs(mean))
C2012 <- summarise_all(filter(centers,Year==2012),funs(mean))
C2013 <- summarise_all(filter(centers,Year==2013),funs(mean))
C2014 <- summarise_all(filter(centers,Year==2014),funs(mean))
C2015 <- summarise_all(filter(centers,Year==2015),funs(mean))
C2016 <- summarise_all(filter(centers,Year==2016),funs(mean))
C2017 <- summarise_all(filter(centers,Year==2017),funs(mean))
center_avgs <- rbind(C2010,C2011,C2012,C2013,C2014,C2015,C2016,C2017)
center_avgs$Pos <- rep("C")

avg_pos <- rbind(center_avgs,pgs_avgs)


#Plots
centers_3PA <- ggplot(centers,aes(x=as.factor(Year),y=X3PA)) + 
  geom_boxplot(color="#485b7a",fill="black",outlier.color = "#dfd8d1") +
  ylab("Three Point Attempts") +
  xlab("Season") +
  ggtitle("Three Point Attempts for Centers") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

pgs_3PA <- ggplot(pgs,aes(x=as.factor(Year),y=X3PA)) +
  geom_boxplot(color="#edcdff",fill="black",outlier.color = "#fff0d5") +
  ylab("Three Point Attempts") +
  xlab("Season") +
  ggtitle("Three Point Attempts for Point Guards") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))
  
pos_mean_three_att <- ggplot(avg_pos,aes(x=interaction(Pos,Year),y=X3PA,fill=Pos)) +
  geom_bar(stat = "identity",color="black") +
  xlab("Interaction of Season and Position") +
  ylab("Average Three Point Attempts") +
  scale_fill_manual(values = c("#c3a5d5","#8560a2")) +
  ggtitle("Average Three Point Attempts per Season of Point Guards and Centers") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "white"),
        plot.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        legend.text = element_text(color = "white"),
        title = element_text(color = "white"))

centers_3PA
pgs_3PA
pos_mean_three_att
```
These plots offer some interesting conclusions about positions in the nba. The bar plot would
make it seem that centers are taking more 3 point attempts than they have in the past, but not at
that high of a rate, as point guards are still taking well over 4 times the amount that centers
are, even this most recent season. However, the box plots show a similar trend among centers as
in team trends. Though the mean amount of 3 point attempts is still very low for centers, there
are many outliers that show individual players taking steps toward a more three point focused
strategy. There are many centers that are attempting more 3 point shots than the average for
point guards over many seasons. The amount of centers doing this has increased over the years as
has the amount of 3 point shots those players are taking. Thought point guards are still making 3
point shots and attempting them at a high rate, centers could begin to challenge the abilites of
smaller players in the coming seasons.

#Part 4
The 3 point shot started in the NBA in 1979, 
so its impact can be seen and measured
based on what happened before and after 1979.
The Golden State warriors became an elite 3-point
shooting team in 2014 and won three NBA championships
since 2014. We will examine the impact of the 3-point
shot and the impact of the rise of the Golden State
Warriors as an elite 3-point shooting team.
```{r part 4.1}
nbaStatsTibble <- as_tibble(nbaStats)
nbaRatingsTibble <- as_tibble(nbaRatings)
# offensive efficiency by years
offensiveRatings <- nbaRatingsTibble %>% group_by(Season) %>% summarize(Rating=mean(ORtg))
# impact the 3-point line had on scoring trends since 
# the Golden State Warriors
# became an elite 3-point shooting team
# and won NBA champships.
scoringAvg <- nbaStatsTibble %>% group_by(Season) %>% summarize(Avg=mean(PTS / G))
# how the 3-point shot changed shot
# selection over the years
threePointPct <- nbaStatsTibble %>% group_by(Season) %>%
  summarize(threePT=mean((X3PA / FGA) * 100))
# graph of the offensive ratings, scoring average 
# and two vs three point shooting averages over the
# years to visualize the data. Keep in mind the
# 3-point line started in 1979 and the Golden
# State Warriors elite 3-point shooting team
# happened in 2014
# offensive ratings plot
p <- ggplot(offensiveRatings, aes(y=Rating,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("Offensive Ratings Trend over the Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
# scoring average
p <- ggplot(scoringAvg, aes(y=Avg,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("Scoring Average Trend over the Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  ylab("Team Scoring Average")

# percentage growth of 3-point shots
p <- ggplot(threePointPct, aes(y=threePT,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("3-Point Shots Taken as a Percent of Overall Shots ") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  ylab("3-Point Percent")
```
When the 3-point shot was introduced
in 1979, we see an immediate jump in offensive 
ratings between 1981 - 1983. After 1983 the
ratings fluctate a bit but remain higher 
compared to the time of only 2-point shots
peaking 1991-92. There is a significant drop
in 1996-97 to the lowset level ever then 
a slow rise. Since the Warriors became an
elite 3-point team in 2014 winning three
NBA titles the offensive ratings has been
on the rise reaching the highest levels ever
in the NBA. Team scoring averages follow a
similar trend, but the highest average was
back in 1991-92 when offensive ratings were at
an all time high. In 1991-92 3-point shooting
was at an all time high corresponding to
the historic scoring and offensive ratings
3-point shots taken percentage continued 
to rise with dips in the early 2000's
and early 2010's. We see a noticeable and
meaningful jump in 2014 reaching an all
time high in 2016-17 as the Warriors
became an elite 3 point shooting team
winnig three championships. We can see teams
attempting to emulate their success driving
all league stats higher. The 3-point line
has made a large impact on the NBA driving
all offensive stats higher overall and teams
like the Warriors have had a lot of success
winning multiple championships because of
the 3 point shot.

The Golden State Warriors became an elite NBA
team around the same time they became the most
prolific NBA 3 point shooting team winning
3 NBA championships since 2014. 
```{r part 4.2}
GSWStatsTibble <- nbaStatsTibble %>% filter(Tm == "GSW" | Tm == "GSW*")
GSWRatingsTibble <- nbaRatingsTibble %>% filter(Tm == "GSW" | Tm == "GSW*")

offensiveRatings <- GSWRatingsTibble %>% group_by(Season) %>% summarize(Rating=mean(ORtg))
scoringAvg <- GSWStatsTibble %>% group_by(Season) %>% summarize(Avg=mean(PTS / G))
threePointPct <- GSWStatsTibble %>% group_by(Season) %>%
  summarize(threePT=mean((X3PA / FGA) * 100))

# offensive ratings plot
p <- ggplot(offensiveRatings, aes(y=Rating,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("GSW Offensive Ratings Trend over the Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))

# scoring average
p <- ggplot(scoringAvg, aes(y=Avg,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("GSW Scoring Average Trend over the Years") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  ylab("Team Scoring Average")

# percentage growth of 3-point shots
p <- ggplot(threePointPct, aes(y=threePT,x=Season, fill=Season))
p + geom_bar(stat = "identity") + 
  ggtitle("GSW 3-Point Shots Taken as a Percent of Overall Shots ") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  ylab("3-Point Percent")


```
Looking at the Warriors stats we can see a steady
and meaningful rise in there offensive ratings,
and scoring average as there number of 3-point 
shots continued on a dramatic rise and a lot of
championship success. Major impact on the rise
of the Warriors.

#Conclusion
From all the data and plots, we can observe a real trend in the modern nba. So many different
stats point to the increased use of the 3 point shot in all teams and positions. We can see how
centers have begun shooting 3 pointers at a much higher rate. This trend that GSW has set has
even changed the way certain positions the game. Upward trends of 3 point percent of shot
attempts show a favoring of the shot. Other teams have begun the climb towards GSW's dominance
and it is undeniable that this modern dynasty is redefining the NBA. More research into specific
data, like more individual players, could help us determine even more effects of this trend.
Either way, GSW has created a unique, powerful and influential dynasty and it's never been a
better time to be a fan.  